Why Real-World Subgraph Isomorphism is Easy, \\ and What This Means for Graph Databases

### 2016-09-23 paper/paper.tex
Write a bit more

    When labels are introduced, as commonly seen in graph databases, additional rich behaviour
    arises.  We show that when only slightly better algorithms are used, all commonly used benchmark
    suites for evaluating graph database systems become trivial, and that existing notions of
    difficulty are based almost entirely upon unnecessarily poor performance from a widely used
    subgraph isomorphism algorithm. With this in mind, we argue that widely researched ``filter /
    verify'' indexing technique used in graph databases is a flawed design based upon a widely
    repeated misunderstanding of what makes subgraph isomorphism hard. We explain why this technique
    would not be beneficial, even with harder instances than those currently used, when paired with
    any reasonable modern subgraph isomorphism algorithm.
\end{abstract}

@@

Throughout, our graphs are undirected, and do not have any loops.  The \emph{order} of a graph is
the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a graph $G$.
The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the same vertex
set as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$ and $w$ are
not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$ vertices, and
an edge between each distinct pair of vertices with independent probability $p$.

A graph may have labels on its vertices, in which case we write $\ell(v)$ for the label of a given
vertex $v$. We write $G(n, p, l)$ for the graph $G(n, p)$ with vertex labels selected uniformly at
random from a set of $l$ different labels.

A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
$T$ (called the \emph{target}) is an injective mapping from $\operatorname{V}(P)$ to
$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. If vertices are labelled,
the mapping must preserve labels, so $\ell(v) = \ell(i(v))$. An \emph{induced subgraph isomorphism}
additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$, then $i(v)$
and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$ for a
non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism. Observe that an
induced isomorphism $i : P \hookrightarrow T$ is a non-induced isomorphism $i : P \rightarrowtail T$
which is also a non-induced isomorphism $\shortoverline{i} : \shortoverline{P} \rightarrowtail
\shortoverline{T}$.

@@

On subsequent rows, we show the average number of search nodes used by the different algorithms. In
general, satisfiable instances are easy, until very close to the phase transition. As we hit the
phase transition and move into the unsatisfiable region, we see complexity increase. Finally, as
we pass through the phase transition and move deeper into the unsatisfiable region, instances become
easier again. This behaviour is largely solver-independent, although VF2 has a larger hard region
than Glasgow or LAD. Thus, although we have moved away from a single control parameter, we still
observe the easy-hard-easy pattern seen in many other \NP-complete problems.

@@

\subsection{Flawed Instances}

@@

Two adjacent differently-coloured vertices.

VF2 does not detect the case that there are insufficient, or even no, labels.

@@

The equation is also used by \citet{DBLP:conf/icde/JiangWYZ07}, who claim that ``usually the
verification time dominates the Query Response Time [s]ince the computational complexity of
$T_{\mathit{iso\_test}}$ is NP-Complete''. They note that \begin{displayquote}``Approximately, the
    value of $T_{\mathit{iso\_test}}$ does not change too much with the difference of query. Thus,
the key to reducing query response time is to minimize the size of the candidate answer
set''.\end{displayquote} This claim becomes understandable when one examines the subgraph matching
algorithm chosen for the verification step \citep{DBLP:journals/jacm/Ullmann76}: as the algorithm
predates techniques like arc-consistent all-different propagation \citep{DBLP:conf/aaai/Regin94}, it
cannot immediately detect unsatisfiability in simple cases like there being two red vertices in the
pattern but only one in the target.

@@

These beliefs are not entirely unfounded: none of the matching algorithms considered in these papers
will immediately detect if a pattern contains two red vertices, whilst the target graph contains
only one. This kind of flaw \emph{should} be picked up at the top of search by an all-different
propagator, but such techniques are not used in graph databases systems, let alone the more powerful
degree- and difference-based propagation rules introduced LAD \cite{Solnon:2010} or the
distance-based reasoning in more recent work \cite{Audemard:2014,McCreesh:2015}.

Indeed, as \citet{DBLP:journals/pvldb/KatsarouNT15} note, some variation on VF2
\citep{Cordella:2004} is the usual matching algorithm of choice, although
\citeS{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes chosen. Other approaches have been
considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08} propose an algorithm which makes use
of the frequency of various features to guide search.  \citet{DBLP:journals/pvldb/LeeHKL12}
determine experimentally that this technique tends to be very effective, even on families of graphs
for which it was not designed.

@@

We do not present the ultimate big data subgraph matching algorithm.

Read outside the field. AI has a lot to offer.

Much more focus on the matching algorithm.

Optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
all-different. Presolving.

Invariants on labels used to inform algorithm design and to provide filtering during search.

Indexing can be used to enhance propagation.

On the other hand, our key message is not just saying use a better algorithm. Need a full
understanding of every aspect of the system, and cannot treat subgraph matching as a black box.

@@

?? Stuff about graph databases.

@@

>>>>

    When labels are introduced, as commonly seen in graph databases, additional rich behaviour
    arises.  We show that when only slightly better algorithms are used, all commonly used benchmark
    suites for evaluating graph database systems become trivial, and that existing notions of
    difficulty are based almost entirely upon unnecessarily poor performance from overly simplistic
    subgraph isomorphism algorithms. With this in mind, we argue that the widely researched ``filter
    / verify'' indexing technique used in graph databases is a flawed design based upon a repeated
    misunderstanding of what makes subgraph isomorphism hard. We explain why this technique would
    not be beneficial, even with harder instances than those currently used, when paired with any
    reasonable modern subgraph isomorphism algorithm.
\end{abstract}

@@

Throughout, our graphs are undirected, and do not have any loops.  The \emph{order} of a graph is
the cardinality of its vertex set. We write $\operatorname{V}(G)$ for the vertex set of a graph $G$.
The \emph{complement} of a graph $G$, denoted $\shortoverline{G}$, is the graph with the same vertex
set as $G$, and with an edge between distinct vertices $v$ and $w$ if and only if $v$ and $w$ are
not adjacent in $G$. We write $G(n, p)$ for an Erd\H{o}s-R\'enyi random graph with $n$ vertices, and
an edge between each distinct pair of vertices with independent probability $p$.  A graph may have
labels on its vertices, in which case we write $\ell(v)$ for the label of a given vertex $v$.

A \emph{non-induced subgraph isomorphism} from a graph $P$ (called the \emph{pattern}) to a graph
$T$ (called the \emph{target}) is an injective mapping $i$ from $\operatorname{V}(P)$ to
$\operatorname{V}(T)$ which preserves adjacency---that is, for every adjacent $v$ and $w$ in
$\operatorname{V}(P)$, the vertices $i(v)$ and $i(w)$ are adjacent in $T$. If vertices are labelled,
the mapping must preserve labels, so $\ell(v) = \ell(i(v))$. An \emph{induced subgraph isomorphism}
additionally preserves non-adjacency---that is, if $v$ and $w$ are not adjacent in $P$, then $i(v)$
and $i(w)$ must not be adjacent in $T$. We use the notation $i : P \rightarrowtail T$ for a
non-induced isomorphism, and $i : P \hookrightarrow T$ for an induced isomorphism. Observe that an
induced isomorphism $i : P \hookrightarrow T$ is a non-induced isomorphism $i : P \rightarrowtail T$
which is also a non-induced isomorphism $\shortoverline{i} : \shortoverline{P} \rightarrowtail
\shortoverline{T}$.

@@

On subsequent rows, we show the average number of search nodes used by the different algorithms. In
general, satisfiable instances are easy, until very close to the phase transition. As we hit the
phase transition and move into the unsatisfiable region, we see complexity increase. Finally, as
we pass through the phase transition and move deeper into the unsatisfiable region, instances become
easier again. This behaviour is largely solver-independent, although VF2 has a larger hard region
than Glasgow or LAD (and the observant reader may notice that VF2 finds some instances extremely
hard when the target is empty but the pattern is not---this turns out to be extremely important, and
we return to it in \cref{section:labelled,section:filterverify}). Thus, although we have moved away
from a single control parameter, we still observe the easy-hard-easy pattern seen in many other
\NP-complete problems.

@@

\subsection{Flawed and Obviously Unsatisfiable Instances}

@@

VF2 does not detect the case that there are insufficient labels. Also empty pattern but
not empty target.

@@

The equation is also used by \citet{DBLP:conf/icde/JiangWYZ07}, who claim that ``usually the
verification time dominates the Query Response Time [s]ince the computational complexity of
$T_{\mathit{iso\_test}}$ is NP-Complete''. They note that \begin{displayquote}``Approximately, the
    value of $T_{\mathit{iso\_test}}$ does not change too much with the difference of query. Thus,
the key to reducing query response time is to minimize the size of the candidate answer
set''.\end{displayquote} This claim becomes understandable when one examines the subgraph matching
algorithm chosen for the verification step \citep{DBLP:journals/jacm/Ullmann76}: as the algorithm
predates techniques like arc-consistent all-different propagation \citep{DBLP:conf/aaai/Regin94}, it
cannot immediately detect unsatisfiability in simple cases like there being two red vertices in the
pattern but only one in the target\footnote{?? expand on: Although interestingly, this algorithm
effectively does forward-checking and decent value ordering heuristics before these concepts
appeared in the constraints literature.}.

@@

These beliefs are not entirely unfounded: none of the matching algorithms considered in these papers
will immediately detect if a pattern contains two red vertices, whilst the target graph contains
only one. This kind of flaw \emph{should} be picked up at the top of search by an all-different
propagator \citep{DBLP:conf/aaai/Regin94}.  However, as \citet{DBLP:journals/pvldb/KatsarouNT15}
note, some variation on VF2 \citep{Cordella:2004} is the usual matching algorithm of choice for
graph database systems, although \citeS{DBLP:journals/jacm/Ullmann76}'s algorithm is sometimes
chosen. Other approaches have been considered: for example, \citet{DBLP:journals/pvldb/ShangZLY08}
propose an algorithm which makes use of the frequency of various features to guide search.
\citet{DBLP:journals/pvldb/LeeHKL12} determine experimentally that this technique tends to be very
effective, even on families of graphs for which it was not designed.

@@

?? The point of this paper is not to present the ultimate big data subgraph matching algorithm.

Read outside the field. AI has a lot to offer. Much more focus on the matching algorithm.

?? Optimisations for trivial cases. Lazy forward checking, cheap alternatives to heuristics and
all-different. Presolving.

?? Invariant which are effective on labels used to inform algorithm design and to provide filtering
during search. Minimum unsat core. Indexing can be used to enhance propagation.

?? On the other hand, our key message is not just saying use a better algorithm. Need a full
understanding of every aspect of the system, and cannot treat subgraph matching as a black box.

@@

@@

?? This is not purely of theoretical interest. Stuff about graph databases.
